---
title: "Codebook wearableComputingDataSet"
author: "Ravi Addanki"
date: "9/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Codebook for wearableComputingDataSet

This is a codebook for tidy data set prepared for wearableComputing project. Please refer to the original code book at http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones  for further details.


## Original Data set information 


The experiments were carried out with a group of 30 volunteers within an age bracket of 19-48 years. They performed a protocol of activities composed of six basic activities: three static postures (standing, sitting, lying) and three dynamic activities (walking, walking downstairs and walking upstairs). The experiment also included postural transitions that occurred between the static postures. These are: stand-to-sit, sit-to-stand, sit-to-lie, lie-to-sit, stand-to-lie, and lie-to-stand. All the participants were wearing a smartphone (Samsung Galaxy S II) on the waist during the experiment execution. We captured 3-axial linear acceleration and 3-axial angular velocity at a constant rate of 50Hz using the embedded accelerometer and gyroscope of the device. The experiments were video-recorded to label the data manually. The obtained dataset was randomly partitioned into two sets, where 70% of the volunteers was selected for generating the training data and 30% the test data.

The sensor signals (accelerometer and gyroscope) were pre-processed by applying noise filters and then sampled in fixed-width sliding windows of 2.56 sec and 50% overlap (128 readings/window). The sensor acceleration signal, which has gravitational and body motion components, was separated using a Butterworth low-pass filter into body acceleration and gravity. The gravitational force is assumed to have only low frequency components, therefore a filter with 0.3 Hz cutoff frequency was used. From each window, a vector of 561 features was obtained by calculating variables from the time and frequency domain. See 'features_info.txt' for more details.

This dataset is an updated version of the UCI Human Activity Recognition Using smartphones Dataset that can be found at: [Web Link]
This version provides the original raw inertial signals from the smartphone sensors, instead of the ones pre-processed into windows which were provided in version 1. This change was done in order to be able to make online tests with the raw data. Moreover, the activity labels were updated in order to include postural transitions that were not part of the previous version of the dataset.

## Additional Variables

   The descriptions of variables can be found in features.txt file in the zipfile of original data set (https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip). The special characters like "(" or ")" or "-" or "," are avoided in variable names per standard practices.
   activityname  - gives the description of activity (standing, sitting, lying,walking, walking downstairs and walking upstairs)
   subject gives the sequence number given to participant. Details of participants are not avaialble, to protect privacy.

Note: activityname can be single value from set (WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING)

Variable                                                          	type	Units
activityname                                                      	Character 
subject                                                           	Integer
TimeDomainBodyAccelerometerMeanX                                  	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerMeanY                                  	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerMeanZ                                  	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerStandardDeviationX                     	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerStandardDeviationY                     	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerStandardDeviationZ                     	numeric	standard gravity units 'g'
TimeDomainGravityAccelerometerMeanX                               	numeric	standard gravity units 'g'
TimeDomainGravityAccelerometerMeanY                               	numeric	standard gravity units 'g'
TimeDomainGravityAccelerometerMeanZ                               	numeric	standard gravity units 'g'
TimeDomainGravityAccelerometerStandardDeviationX                  	numeric	standard gravity units 'g'
TimeDomainGravityAccelerometerStandardDeviationY                  	numeric	standard gravity units 'g'
TimeDomainGravityAccelerometerStandardDeviationZ                  	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerJerkMeanX                              	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerJerkMeanY                              	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerJerkMeanZ                              	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerJerkStandardDeviationX                 	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerJerkStandardDeviationY                 	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerJerkStandardDeviationZ                 	numeric	standard gravity units 'g'
TimeDomainBodyGyroscopeMeanX                                      	numeric	radians/sec
TimeDomainBodyGyroscopeMeanY                                      	numeric	radians/sec
TimeDomainBodyGyroscopeMeanZ                                      	numeric	radians/sec
TimeDomainBodyGyroscopeStandardDeviationX                         	numeric	radians/sec
TimeDomainBodyGyroscopeStandardDeviationY                         	numeric	radians/sec
TimeDomainBodyGyroscopeStandardDeviationZ                         	numeric	radians/sec
TimeDomainBodyGyroscopeJerkMeanX                                  	numeric	radians/sec
TimeDomainBodyGyroscopeJerkMeanY                                  	numeric	radians/sec
TimeDomainBodyGyroscopeJerkMeanZ                                  	numeric	radians/sec
TimeDomainBodyGyroscopeJerkStandardDeviationX                     	numeric	radians/sec
TimeDomainBodyGyroscopeJerkStandardDeviationY                     	numeric	radians/sec
TimeDomainBodyGyroscopeJerkStandardDeviationZ                     	numeric	radians/sec
TimeDomainBodyAccelerometerMagnitudeMean                          	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerMagnitudeStandardDeviation             	numeric	standard gravity units 'g'
TimeDomainGravityAccelerometerMagnitudeMean                       	numeric	standard gravity units 'g'
TimeDomainGravityAccelerometerMagnitudeStandardDeviation          	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerJerkMagnitudeMean                      	numeric	standard gravity units 'g'
TimeDomainBodyAccelerometerJerkMagnitudeStandardDeviation         	numeric	standard gravity units 'g'
TimeDomainBodyGyroscopeMagnitudeMean                              	numeric	radians/sec
TimeDomainBodyGyroscopeMagnitudeStandardDeviation                 	numeric	radians/sec
TimeDomainBodyGyroscopeJerkMagnitudeMean                          	numeric	radians/sec
TimeDomainBodyGyroscopeJerkMagnitudeStandardDeviation             	numeric	radians/sec
FrequencyDomainBodyAccelerometerMeanX                             	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerMeanY                             	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerMeanZ                             	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerStandardDeviationX                	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerStandardDeviationY                	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerStandardDeviationZ                	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerMeanFreqX                         	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerMeanFreqY                         	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerMeanFreqZ                         	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkMeanX                         	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkMeanY                         	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkMeanZ                         	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkStandardDeviationX            	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkStandardDeviationY            	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkStandardDeviationZ            	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkMeanFreqX                     	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkMeanFreqY                     	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerJerkMeanFreqZ                     	numeric	standard gravity units 'g'
FrequencyDomainBodyGyroscopeMeanX                                 	numeric	radians/sec
FrequencyDomainBodyGyroscopeMeanY                                 	numeric	radians/sec
FrequencyDomainBodyGyroscopeMeanZ                                 	numeric	radians/sec
FrequencyDomainBodyGyroscopeStandardDeviationX                    	numeric	radians/sec
FrequencyDomainBodyGyroscopeStandardDeviationY                    	numeric	radians/sec
FrequencyDomainBodyGyroscopeStandardDeviationZ                    	numeric	radians/sec
FrequencyDomainBodyGyroscopeMeanFreqX                             	numeric	radians/sec
FrequencyDomainBodyGyroscopeMeanFreqY                             	numeric	radians/sec
FrequencyDomainBodyGyroscopeMeanFreqZ                             	numeric	radians/sec
FrequencyDomainBodyAccelerometerMagnitudeMean                     	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerMagnitudeStandardDeviation        	numeric	standard gravity units 'g'
FrequencyDomainBodyAccelerometerMagnitudeMeanFreq                 	numeric	standard gravity units 'g'
FrequencyDomainBodyBodyAccelerometerJerkMagnitudeMean             	numeric	standard gravity units 'g'
FrequencyDomainBodyBodyAccelerometerJerkMagnitudeStandardDeviation	numeric	standard gravity units 'g'
FrequencyDomainBodyBodyAccelerometerJerkMagnitudeMeanFreq         	numeric	standard gravity units 'g'
FrequencyDomainBodyBodyGyroscopeMagnitudeMean                     	numeric	radians/sec
FrequencyDomainBodyBodyGyroscopeMagnitudeStandardDeviation        	numeric	radians/sec
FrequencyDomainBodyBodyGyroscopeMagnitudeMeanFreq                 	numeric	radians/sec
FrequencyDomainBodyBodyGyroscopeJerkMagnitudeMean                 	numeric	radians/sec
FrequencyDomainBodyBodyGyroscopeJerkMagnitudeStandardDeviation    	numeric	radians/sec
FrequencyDomainBodyBodyGyroscopeJerkMagnitudeMeanFreq             	numeric	radians/sec




## Prepared tidy data set
 You may find the prepared tidy data set at "./output1/tidyDataSet.txt" 

```{r tidyDataset}
tidyDataset <-read.table("./output1/tidyDataSet.txt",header = TRUE)
summary(tidyDataset)
```
